Measurement of dark matter halo concentrations using a bayesian approach

### 2014-05-23 paper/article.tex
adding text

\section{Basics of the NFW profile}
\label{sec:basics}

The Navarro-Frenk-White mass profile can be normalized respect to the virial mass and written in function of the normalized radius respect to the virial radius $r=R/R_{vir}$ and the concentration $c$ 
\[
m_{nfw}\left(r;c\right)=A\left[\log\left(1+rc\right)-\left(\frac{rc}{rc+1}\right)\right]
\]
Where
\[
A=\left[\log\left(1+c\right)-\left(\frac{c}{c+1}\right)\right]^{-1} 
\]
\section{Bayesian Method}
\label{sec:method}

In order to obtain a mass profile for each halo in funcion of the radius, the center of each halo must be calculated. However the center of mass will not give an accurate position for the center because some haloes are highly irregular, is more convenient to take the gravitational potential for each particle
\[
\phi\left(\boldsymbol{R}_{n}\right)=-\sum_{j\neq n}\frac{1}{\left\Vert \boldsymbol{R}_{n}-\boldsymbol{R}_{j}\right\Vert }
\]
Take $\boldsymbol{R}_{0}=\min\left(\phi\left(\boldsymbol{R}_{n}\right)\right)$ as the center then define new coordinates $\boldsymbol{R}_{n}^{\prime}=\boldsymbol{R}_{0}-\boldsymbol{R}_{n}$, this $\boldsymbol{R}_{0}$ will be in the most relevant region for the dynamics of each halo. Organizing the $\left\{ \boldsymbol{R}_{n}^{\prime}\right\} $ in crescent order, the accumulated mass for each $\boldsymbol{R}_{n}^{\prime}$ will be $M_{n}=nM$.

\begin{figure}[H]
\caption{FIXME: Mass profile of a halo}
\end{figure}

Then the virial radius is obtained and the particles that are beyond the virial radius are removed. After that both the mass and the radius get normalized respect to the mass and virial radius, we will call these new variables $m_{n}$ and $r_{n}$. The last part of the process consists in fitting the normalized NFW mass profile to the $m_{n}$ data using the Metropolis-Hastings algorithm to sample the Likelihood ${\cal L}\left(c\right)=\exp\left(-\chi^{2}\left(c\right)\right)$ where 
\[
\chi^{2}\left(c\right)=\sum\limits _{n}\left|m_{n}-m_{nfw}\left(r_{n},c\right)\right|^{2}
\]
And taking the values of $c$ where the Likelihood is maximum. Moreover, we can obtain a confidence interval for $c$ doing an histogram of the random walk done by the Metropolis-Hastings Algorithm. Finally we compare this results with the values obtaned by BDM, since the number of halos obtained by FOF and BDM is different, each BDM halo gets related with the nearest halo in FOF. We compute two concentrations, one where the overdensity limit is $360\rho_{back}$ (like in BDMV) an other where the overdensity limit is $740\rho_{back}$ (like in BDMW).

\section{Mock Data and Simulations}
\label{sec:data}
\section{Results}
\label{sec:results}

>>>>

\section{Basic properties of the NFW density profile}
\label{sec:basics}

The Navarro-Frenk-White density profile can be written as

\begin{equation}
\rho(r) = \frac{\rho_c\delta_c}{r/r_s(1+r/r_s)^2}, 
\end{equation}
%
where $\rho_c\equiv 3H^2/8\pi G$ is the Universe critical density,
$\delta_c$ is the halo dimensionless characteristic density and $r_s$
is known as the scale radius, the radius that marks the transition
between the two power law behaviour in the $\rho\propto r^{-1}$ for
$r<r_s$ and $\rho\propto r^{-2}$ for  $r>r_s$.

We define the virial radius of a halo, $r_v$, as the boundary of the
spherical volume that encloses an average density of $\Delta_v$ times
the critical density of the Universe. The corresponding mass $M_{v}$
can be thus expressed as $M_{v} = \frac{4\pi}{3}\rho_c\Delta_v r_v^3$.

The total mass enclosed within a radius $r$ can be computed to be:
\begin{equation}
M(<r) = 4\pi\rho_c\delta_c  r_s^3\left[\ln \left
  (\frac{r_s+r}{r}\right) - \frac{r}{r_s+r}\right].
\end{equation}

We can now define the concentration of the halo as $c=r_v/r_s$, the
dimensioless variable $x\equiv r/r_v$ and $m\equiv M(<r)/M_v$, which
allows us to express the total enclosed mass within a dimensionless
radius $x$ as:

\begin{equation}
m(<x) =
\frac{1}{A}\left[\ln\left(1+xc\right)-\left(\frac{xc}{xc+1}\right)\right],
\label{eq:profile}
\end{equation}
%
where 
%
\begin{equation}
A=\left[\ln\left(1+c\right)-\left(\frac{c}{c+1}\right)\right],
\end{equation}

meaning that the concentration is the only free parameter to
determine the density profile of the halo.

\begin{figure}
\caption{FIXME: Mass profile of a halo}
\end{figure}

\section{A Bayesian Approach to Halo Fitting}
\label{sec:method}

We proceed to find the value of the concentration parameter that
best describes the simulation data, following the model in
Eq. (\ref{eq:profile}). We use the integrated mass profile because it
allows us to use the data directly with the simulation without binning
the particle positions and estimating a density.

We construct the integrated mass profile by ranking the particles by
their increasing distance to the center of the halo. Once they are ranked,
the total mass at a radius $r_i$, increases by $m_p$, where $r_i$ is
the position of the $i$-th particle and $m_p$ is the mass of the
computational particle.  We define the center of the halo to be  at
the position of the particle with the lowest gravitational
potential. In the process of building the mass profile we discard the
particle at the center.

We stop the construction of the integrated mass profile once we arrive
at an average density of $\Delta_h\rho_c$. This radius marks the
virial radius and the virial mass. We divide the total mass enclosed
mass $M_i$ and the radii $r_i$ by these values to obtaine the
dimensionless variables $x_i$ and $m_i$.

Using these positions and masses we define the following $\chi^2$ function

\begin{equation}
\chi^2(c) = \sum_{i}[m_i - m(< x_i;c)]^2, 
\end{equation}
%
where $m(<x_i;c)$ corresponds to the values in Eq.(\ref{eq:profile}) at
$x=x_i$ and a given value of the concentration parameter $c$.

Finally we use a Metropolis-Hastings algorithm to sample the likelihood
function distribution defined by ${\cal L}(c)=\exp(-\chi^2(c)/2)$ to
find the optimum value of $c$ and its associated uncertainty
$\sigma_c$.

\section{Numerical Experiments}
Finally we compare this results with the values obtaned by BDM, since
the number of halos obtained by FOF and BDM is different, each BDM
halo gets related with the nearest halo in FOF. We compute two
concentrations, one where the overdensity limit is $360\rho_{back}$
(like in BDMV) an other where the overdensity limit is
$740\rho_{back}$ (like in BDMW).

\subsection{Mock Data and Simulations}
\label{sec:data}
\section{Results}
\label{sec:results}
